### 2024-11-08 MetadataS1/MetadataS1.tex
update manuscrit data paper open research statement and minor changes lapland longspur and Baird sandpiper

\begingroup  
  \centering
\textbf{Long-term abundance time-series of the High Arctic terrestrial vertebrate community of Bylot Island, Nunavut}\\[1.5em]
 Louis Moisan\textsuperscript{1,2}, Azenor Bideault\textsuperscript{2,3}, Gilles Gauthier\textsuperscript{4}, Éliane Duchesne\textsuperscript{1}, 
 Dominique Fauteux\textsuperscript{5}, Dominique Berteaux\textsuperscript{1}, Pierre Legagneux\textsuperscript{3,6}, Marie-Christine Cadieux \textsuperscript{3} and Joël Bêty\textsuperscript{1}\\[1.5em]
\textbf{Author Affiliations}:\\
\textsuperscript{1} Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Chimie et Géographie, Université du Québec à Rimouski, Rimouski, QC, Canada\\
\textsuperscript{2} Chaire de Recherche du Canada en Écologie Intégrative, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Université de Sherbrooke, Sherbrooke, QC, Canada\\
\textsuperscript{3} Chaire de Recherche Sentinelle Nord sur l’Impact des Migrations Animales sur les Écosystèmes Nordiques, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Université Laval, Québec, QC, Canada\\
\textsuperscript{4} Centre d’Études Nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
\textsuperscript{5} Centre d’Études Nordiques, Centre de Connaissance et d’Exploration de l’Arctique, Musée Canadien de la Nature, Ottawa, ON, Canada\\
\textsuperscript{6} Centre d’Études Biologiques de Chizé (CEBC-CNRS), Université de La Rochelle, France 
\\[1.5em]
\textbf{Corresponding Author}:\\
\textit{\href{mailto:louis.moisan.bio@gmail.com}{louis.moisan.bio@gmail.com}}\\
[1.5em]
\textbf{Open Research statement}:\\
The complete data set is available as Supporting Information at: \url{https://doi.org/10.5061/dryad.44j0zpcnt}. Metadata document, raw data and codes are also available at \url{https://doi.org/10.5061/dryad.44j0zpcnt}.\\
\endgroup
\newpage

@@

        \item[] \textit{\textbf{Long-tailed jaeger and parasitic jaeger}} \\
        Since 2004, observers have walked parallel transects spaced 400 meters apart, covering the entire southern side of the glacial river in the Qarlikturvik Valley (33 km\textsuperscript{2}; \textbf{Figure \ref{figure:qarlikturvik_valley}}), during the nesting period. The aim of those transects was to record nests of long-tailed jaegers, parasitic jaegers, and sandhill cranes. Observers listened for alarm calls to detect territorial birds, and then located nests by observing the birds returning to their nests from elevated vantage points. We consider the sampling to be systematic for long-tailed and parasitic jaeger, since those species tend to leave their nest relatively far from the observer to perform mobbing behavior, and thus increasing their detection probability. We do not consider the sampling to be systematic for sandhill cranes as they only display defensive behaviors near their nests at relatively short distances (see opportunistic nest monitoring below).

@@

\subsubsection*{c. Observation of individuals}
\item[] \textit{\textbf{Vertebrate count transects}}\\
From 2010 to 2023, observers walked 500-meter linear transects where all vertebrate individuals observed within 150 meters on either side were counted (146 to 320 transects per year). Transects were distributed across all lowland zones of the study area, typically in mesic habitat, and were carried out during the nesting period (between June 21 and July 14; \citet{duchesne2021}; \textbf{Figure \ref{figure:transects}}). Furthermore, specifically for American golden-plovers, we measured the distance of each observed individual to the transect path.
\\
  \includegraphics[width=0.7\textwidth, angle=0]{figures/transects.pdf}
  \vspace{-35pt} % Reduce space between figure and caption
  \caption{Spatial distribution of vertebrate count transects (white lines) on the south plain of Bylot Island (146 to 320 transects per year).}
 \label{figure:transects}
\item[] \textit{\textbf{Snow goose point count}}\\
At the start, middle, and end of each vertebrate count transect, a point count with a radius of 125 meters was conducted to determine the number of snow goose breeding pairs. On average, 613 \textpm{142} point counts were sampled each year, covering an area of 30 \textpm{7} km\textsuperscript{2}.
\\
\item[] \textit{\textbf{Incidental observations}}\\
Since 2007, observers have recorded all vertebrate species observed opportunistically during field work and tallied the total number of individuals at the end of each day \citep{gauthier2020daily, gauthier2024a}. The number of hours spent in the field served as a proxy for the sampling effort. We used the number of individuals observed per hour spent in the field calculated by \citet{gauthier2024a} as an index of relative abundance for each species. Moreover, we separated observations made in lowland from those in upland zones to have a relative abundance of each species in each of these two broad categories (\textbf{Table \ref{table:table_species_relative_abundance_upland}}). Given that incidental observations lacked georeferencing, we opted to extract upland observations by focusing on observations made during visits to rough-legged hawk nests, which are mostly located in uplands. \\
\input{tables/table_species_relative_abundance_upland.tex}
%label-> table_species_relative_abundance_upland
\item[] \textit{\textbf{Testimonials of ermine sightings}}\\
There was no direct estimation of ermine abundance on Bylot Island as they are quite difficult to obtain. The density estimates for ermine were derived from an annual abundance index established by \citet{bolduc2023}, which relied on testimonials provided by observers across the whole study area from 1993 to 2019. The testimonials provided by observers were used to create an abundance index ranging from 0 to 3 \citep{bolduc2023}. In this index, a score of 0 corresponds to the absence of ermine sightings, 1 indicates a single sighting of a lone individual, 2 represents multiple sightings of lone individuals, and 3 signifies at least one sighting of a family group. Scores of individual participants were averaged annually as detailed in \citet{bolduc2023}. 
\end{enumerate}
\subsubsection*{d. Capture of individuals}
\item[] \textit{\textbf{Lemming trapping}}\\
Since 2004, Nearctic brown and collared lemmings were live-trapped 3 times during the summer (mid-June, mid-July, and mid-August) in two 11 ha grids. Each grid is made of 144 traps separated by 30 m according to a cartesian plane, one in mesic habitat and the other in wet habitat, located in the Qarlikturvik valley \citep{fauteux2015, gauthier2020lemmings}. Density of each species was estimated at each occasion using spatially explicit capture-recapture methods (see \citet{fauteux2015} for details). From 1995 to 2016 snap-trapping was performed once a year (mid-July) along 2 groups of transects located in the same habitats than the trapping grids \citep{gruyer2008}. Index of abundance derived from snap-trapping were transformed in density estimates in each habitat for the period 1995-2003 using the equation provided by \citet{fauteux2018} based on the period of overlap between the two sampling methods (2004 to 2016). 
\\
\item[] \textit{\textbf{Arctic fox movement tracking}}\\
In order to assess fox abundance based on the size of their home range, 109 Arctic foxes were fitted with Argos Platform Transmitter Terminals mounted on collars between 2008 and 2016 \citep{lai2015,christin2015, dulude2023}. Foxes were captured between May and August across the study area, within and outside the goose colony \citep{dulude2023}. Sampling of animal locations was set for an interval of 1 or 2 days and only locations between May 1 and October 30 were retained \citep{dulude2023}.
\\
\item[] \textit{\textbf{Parasitic jaeger banding}}\\
In 2009, a significant effort was made to band as many parasitic jaegers as possible within the study area. This effort resulted in the banding of 17 adult individuals (Therrien and Gauthier, unpublished data).
\end{enumerate}
\subsubsection*{e. Species body mass}
All vertebrate individuals captured for marking purposes were systematically weighed: snow goose (G. Gauthier, M.-C. Cadieux and J. Lefebvre, unpublished data), snowy owl \citep{therrien2012, robillard2018}, American-golden plovers \citep{lamarre2021}, common-ringed plovers \citep{leandri2019}, other shorebirds (J. Bêty, unpublished data), glaucous gulls \citep{gauthier2015}, long-tailed jaeger \citep{seyer2019}, parasitic jaegers (J.-F. Therrien and G. Gauthier, unpublished data), Lapland longspurs (J. Bêty and G. Gauthier, unpublished data), lemmings \citep{gauthier2020lemmings}, American ermine (Bilodeau and Bolduc, unpublished data) and Arctic foxes \citep{lai2015}. When not available, we extracted mean body mass from the literature \citep{wilman2014}.

@@

        \begin{enumerate}[label=\alph*.]       
     \item[] \textit{\textbf{Pacific loon, red-throated loon, cackling goose, tundra swan and glaucous gull}}   \\
Based on the systematic and intensive search for the glaucous gull, cackling goose, tundra swan, red- throated loon and Pacific loon nests in wetlands, we are confident that we have found nearly all nests across the study area from 2017 to 2019 and in 2022. We observed a relatively strong correlation between the nest density of glaucous gulls in the Qarlikturvik valley and the nest density across the entire study area (R\textsuperscript{2}= 0.84, p = 0.16, n= 4). Consequently, we estimated the density of glaucous gulls at the scale of the study area between 2004 and 2016 based on the nest density in the Qarlikturvik valley ($y = 0.12409x + 0.13774$). However, we did not observe such strong relationships for loons and swans and thus we did not extend the time series. Regarding cackling geese, we observed signs of an exponential increase over time based on the annual number of nests found in various zones of the study area. We thus fitted an exponential model using the number of nests found annually over two distinct periods: in 1996 when the first nest was discovered, and then from 2017 to 2023 when sampling effort was systematic across the whole study area (\textbf{Figure \ref{figure:cackling}}). We used the fitted model to estimate abundance between 1996 and 2016 when monitoring was less systematic, which could potentially underestimate observed abundance as seen on \textbf{Figure \ref{figure:cackling}}. We multiplied nest density by two to obtain the abundance (assuming two individuals per nest).

@@

                \item[] \textit{\textbf{Rough-legged hawk, peregrine falcon and snowy owl}} \\
                We estimated the abundance of breeding rough-legged hawks, peregrine falcons and snowy owls based on systematic nest monitoring conducted throughout the study area for these species. To convert the number of nests into breeding abundance, we multiplied it by two (assuming two individuals per nest). For snowy owls, we extended the time series from 1996 to 2011 based on a linear regression between nest density in the Qarlikturvik valley and nearby plateaus (Black and Southern plateaus) and nest density across the entire study area ($y= 0.68867x -0.00173$; R\textsuperscript{2}= 0.99; p < 0.0001, n= 10). We used the same approach for rough-legged hawks ($y=0.49851x$, R\textsuperscript{2}= 0.99, p < 0.0001, n= 8) to extend the time series from 2007 to 2012. We did not extend the time series for peregrine falcons because the correlation is not as strong (R\textsuperscript{2}= 0.44, p= 0.27, n=8).

@@

                \item[] \textit{\textbf{Sandhill crane}}\\
                We estimated the mean abundance of sandhill cranes in the lowland zones of the study area based on a regression between nest density and the number of individuals observed per transect (\textbf{Figure \ref{figure:sacr}}). In this relationship, nest density and transect observations come from the 8 km\textsuperscript{2} area of the Qarlikturvik valley where extensive nest search is performed. We acknowledge that the opportunistic monitoring of this species likely underestimated the true nest density. However, considering the extensive sampling effort deployed annually within this area, we are confident in obtaining a realistic order of magnitude for the number of nests present. Number of individuals observed along transects in each lowland zone was converted into nest density using the regressions, and then in total number of individuals in each zone by multiplying by the area of the zone and a factor 2. We estimated the density in the upland zones by applying a correction factor to the annual mean density in lowland zones. This correction factor was based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).

\begin{figure}[H]
\centering
\includegraphics[width=0.5\linewidth]{figures/sandhill_crane_nest_transects.pdf} 
\caption{Linear regression between the nest density of sandhill cranes and the number of individuals observed per transect (nest density = 1.12 x number of individuals per transect; regression was forced to pass through the origin). The fit (R\textsuperscript{2} and p value) of the regression (red line) with the empirical data from the Qarlikturvik valley is presented (blue dots). Data points represent annual values.}
\label{figure:sacr}
\end{figure}
                \item[] \textit{\textbf{American golden-plover and black-bellied plover}}\\
                We used a distance sampling approach to estimate the abundance of American golden-plovers in the lowland zones of the study area between 2014 and 2023. Observations of plovers were made along vertebrate count transects mainly in mesic habitat. Perpendicular distance between detected individuals and the transect path were used (n= 1015) to estimate a detection function with the \textit{ds} function from the \textit{Distance} package \citep{miller2019}. To determine the detection function, we applied a truncation distance of 150 m (i.e., maximum distance on either side of the observer where observations have been considered) and selected the model with the lowest AIC, which included a "hn" key and a single "cos" adjustment term. We excluded observations of groups with more than four individuals, as these likely indicated groups of non-breeders. We did not estimate abundance in wetland habitat because American golden-plovers nest almost exclusively in mesic habitat \citep{parmelee1967}. We estimated the abundance in the upland zones (i.e., plateaus) by applying a correction factor to the abundance in lowland zones. This correction factor was based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).

@@

                \item[] \textit{\textbf{Lapland longspur and Baird’s sandpiper}}\\
                We estimated the mean abundance of Lapland longspur in the different lowland zones of the study area based on a regression between nest density and the number of individuals observed per transect (\textbf{Figure \ref{figure:lalo_basa}}). For Baird’s sandpiper, we employed a similar approach, but instead of using the mean number of individuals observed per transect, we used the mean proportion of transects where at least one individual was detected. We made this adjustment because this species was less frequently observed. In this relationship, nest density for these two species came from the intensive nest sampling conducted within the core 2 km\textsuperscript{2} area of the Qarlikturvik valley and observations of individuals from transects carried out in the larger 8 km\textsuperscript{2} area in which the core area was located. This approach allowed us to incorporate a larger sample size from the transects while focusing on a measure of nest density determined systematically. Transects observations in lowland were then converted into nest density using the regressions, and then in total number of individuals by multiplying by the area and a factor 2. We estimated the density of both species in the upland zones by applying a correction factor to the annual mean density in lowland zones. This correction factor was based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}). We acknowledge that the regression for Baird’s sandpiper is weak; however, it offers some refinement compared to assuming a uniform density throughout the study area.

                \begin{figure}[H]
                \begin{subfigure}{0.5\textwidth}
                \includegraphics[width=0.95\linewidth]{figures/lapland_longspur_nest_transect.pdf} 
                \caption{}
                \end{subfigure}
                \begin{subfigure}{0.5\textwidth}
                \includegraphics[width=0.95\linewidth]{figures/bairds_sandpiper_nest_transect.pdf}
                \caption{}
                \end{subfigure}
                \caption{a) Linear regression between the nest density of Lapland longspurs and the number of individuals observed per transect (nest density= 2.3422 x number of individuals per transect; regression was forced through the origin). The fit (R\textsuperscript{2} and p value) of the regression (red line) with the empirical data from the Qarlikturvik valley is presented (blue dots). Data points represent annual values. b) Linear regression between the nest density of Baird’s sandpiper and the proportion of transect with at least one individual observed (nest density= 34.9248 x proportion of transects with at least one individual; regression was forced through the origin). The fit (R\textsuperscript{2} and p value) of the regression (red line) with the empirical data from the Qarlikturvik valley is presented (blue dots). Data points represent annual values.}
                \label{figure:lalo_basa}
                \end{figure}

@@

                \subsubsection*{d. Project personnel}
                \begin{enumerate}[label=\alph*.]
                        \item[] \textit{\textbf{Principal and associated investigators}}\\
                        Gilles Gauthier, Austin Reed, Jean-François Giroux, Dominique Berteaux, Joël Bêty, Josée Lefebvre, Dominique Gravel, Jean-François Therrien, Nicolas Lecomte, Dominique Fauteux, Pierre Legagneux (see \citet{gauthier2024b})
                        \item[] \textit{\textbf{Students}}\\
                        By combining animal and plant ecology, 24 doctoral theses and 56 master theses have been completed in relation to the study area located on the south plain of Bylot Island (see \citet{gauthier2024b}).
                \end{enumerate}
\newpage 
 \section*{\textcolor{Blue}{Class III. Data set status and accessibility}}
    \section*{A. Status}
                \subsection*{1. Latest update} 25 October 2024
        \subsection*{2. Latest archive date} 25 October 2024
        \subsection*{3. Metadata status} 25 October 2024
        \subsection*{4. Data verification} The methods employed to estimate species abundance were subject to several rounds of revision by the authors.
   \section*{B. Accessibility}
                \subsection*{1. Storage location and medium}
The complete data set is available as Supporting Information at: \url{https://doi.org/10.5061/dryad.44j0zpcnt}.\\
Metadata document, raw data and codes are also available at \url{https://doi.org/10.5061/dryad.44j0zpcnt}.

@@

                        \subsection*{2. Size} 
                        a. 35 records, not including header row (4.3 kB)\\
                        b. 271 records, not including header row  (33.3 kB)\\
                        c. 53 records, not including header row  (3.7 kB)\\
                        d. 18 records, not including header row  (961 B)

@@

                        \subsection*{7. Authentication procedures}
                        Sums of the numeric columns: \\
                        b. \texttt{BYLOT-species\_abundance.csv}: abundance= 2280165\\
                        c. \texttt{BYLOT-species\_body\_mass.csv}: body\_mass\_g= 49617; sample\_size= 13902\\
                        d. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}: mean\_nest\_density\_km2= 19.991; sd\_nest\_density\_km2= 10.539; sample\_size\_nest\_density\_km2= 185

@@

 \section*{\textcolor{Blue}{Class V. Supplemental descriptors}}
    \section*{A. Data acquisition}
    		\subsection*{1. Data forms or acquisition methods}
   		See Section \textbf{2. Experimental or sampling design} 
   		\subsection*{2. Location of completed data forms}
        Raw data and codes used to extract the data set and the current document are publicly available at \url{https://doi.org/10.5061/dryad.44j0zpcnt}.
        \subsection*{3. Data entry verification procedures}
       	The methods used to extract final species abundance estimates were subject to several rounds of revision by the authors.

>>>>

\begingroup  
  \centering
\textbf{Long-term abundance time-series of the High Arctic terrestrial vertebrate community of Bylot Island, Nunavut}\\[1.5em]
 Louis Moisan\textsuperscript{1,2}, Azenor Bideault\textsuperscript{2,3}, Gilles Gauthier\textsuperscript{4}, Éliane Duchesne\textsuperscript{1}, 
 Dominique Fauteux\textsuperscript{5}, Dominique Berteaux\textsuperscript{1}, Pierre Legagneux\textsuperscript{3,6}, Marie-Christine Cadieux \textsuperscript{3} and Joël Bêty\textsuperscript{1}\\[1.5em]
\textbf{Author Affiliations}:\\
\textsuperscript{1} Chaire de Recherche du Canada en Biodiversité Nordique, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Chimie et Géographie, Université du Québec à Rimouski, Rimouski, QC, Canada\\
\textsuperscript{2} Chaire de Recherche du Canada en Écologie Intégrative, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Université de Sherbrooke, Sherbrooke, QC, Canada\\
\textsuperscript{3} Chaire de Recherche Sentinelle Nord sur l’Impact des Migrations Animales sur les Écosystèmes Nordiques, Centre d’Études Nordiques, Centre de la Science de la Biodiversité du Québec, Département de Biologie, Université Laval, Québec, QC, Canada\\
\textsuperscript{4} Centre d’Études Nordiques, Département de Biologie, Université Laval, Québec, QC, Canada\\
\textsuperscript{5} Centre d’Études Nordiques, Centre de Connaissance et d’Exploration de l’Arctique, Musée Canadien de la Nature, Ottawa, ON, Canada\\
\textsuperscript{6} Centre d’Études Biologiques de Chizé (CEBC-CNRS), Université de La Rochelle, France 
\\[1.5em]
\textbf{Corresponding Author}:\\
\textit{\href{mailto:louis.moisan.bio@gmail.com}{louis.moisan.bio@gmail.com}}\\
[1.5em]
\textbf{Open Research statement}:\\
The complete data set will be published as supporting information should the manuscript be accepted for publication. Metadata document, raw data and codes will also be archived in Dryad.\\
\endgroup
\newpage

@@

        \item[] \textit{\textbf{Long-tailed jaeger, parasitic jaeger and sandhill cranes}} \\
        Since 2004, observers have walked parallel transects spaced 400 meters apart, covering the entire southern side of the glacial river in the Qarlikturvik Valley (33 km\textsuperscript{2}; \textbf{Figure \ref{figure:qarlikturvik_valley}}), during the nesting period. The aim of those transects was to record nests of long-tailed jaegers, parasitic jaegers, and sandhill cranes. Observers listened for alarm calls to detect territorial birds, and then located nests by observing the birds returning to their nests from elevated vantage points. We consider the sampling to be systematic for long-tailed and parasitic jaeger, since those species tend to leave their nest relatively far from the observer to perform mobbing behavior, and thus increasing their detection probability. We consider the sampling to be opportunistic for sandhill cranes, as individuals display defensive behaviors only at relatively close distances from their nests (see opportunistic nest monitoring below.

@@

\newpage
\subsubsection*{c. Observation of individuals}
\begin{enumerate}[label=\roman*]
\item[] \textit{\textbf{Vertebrate count transects}}\\
From 2010 to 2023, observers walked 500-meter linear transects where all vertebrate individuals observed within 150 meters on either side were counted (146 to 320 transects per year). Transects were distributed across all lowland zones of the study area, typically in mesic habitat, and were carried out during the nesting period (between June 21 and July 14; \citet{duchesne2021}; \textbf{Figure \ref{figure:transects}}). Furthermore, specifically for American golden-plovers, we measured the distance of each observed individual to the transect path.
\\
\begin{figure}
\centering
  \includegraphics[width=0.7\textwidth, angle=0]{figures/transects.pdf}
  \vspace{-35pt} % Reduce space between figure and caption
  \caption{Spatial distribution of vertebrate count transects (white lines) on the south plain of Bylot Island (146 to 320 transects per year).}
 \label{figure:transects}
\end{figure}
\newpage
\item[] \textit{\textbf{Snow goose point count}}\\
At the start, middle, and end of each vertebrate count transect, a point count with a radius of 125 meters was conducted to determine the number of snow goose breeding pairs. On average, 613 \textpm{142} point counts were sampled each year, covering an area of 30 \textpm{7} km\textsuperscript{2}.
\\
\item[] \textit{\textbf{Incidental observations}}\\
Since 2007, observers have recorded all vertebrate species observed opportunistically during field work and tallied the total number of individuals at the end of each day \citep{gauthier2020daily, gauthier2024a}. The number of hours spent in the field served as a proxy for the sampling effort. We used the number of individuals observed per hour spent in the field calculated by \citet{gauthier2024a} as an index of relative abundance for each species. Moreover, we separated observations made in lowland from those in upland zones to have a relative abundance of each species in each of these two broad categories (\textbf{Table \ref{table:table_species_relative_abundance_upland}}). Given that incidental observations lacked georeferencing, we opted to extract upland observations by focusing on observations made during visits to rough-legged hawk nests, which are mostly located in uplands. \\
\input{tables/table_species_relative_abundance_upland.tex}
%label-> table_species_relative_abundance_upland
\newpage
\item[] \textit{\textbf{Testimonials of ermine sightings}}\\
There was no direct estimation of ermine abundance on Bylot Island as they are quite difficult to obtain. The density estimates for ermine were derived from an annual abundance index established by \citet{bolduc2023}, which relied on testimonials provided by observers across the whole study area from 1993 to 2019. The testimonials provided by observers were used to create an abundance index ranging from 0 to 3 \citep{bolduc2023}. In this index, a score of 0 corresponds to the absence of ermine sightings, 1 indicates a single sighting of a lone individual, 2 represents multiple sightings of lone individuals, and 3 signifies at least one sighting of a family group. Scores of individual participants were averaged annually as detailed in \citet{bolduc2023}.

\end{enumerate}
\subsubsection*{d. Capture of individuals}
\begin{enumerate}[label=\roman*]
\item[] \textit{\textbf{Lemming trapping}}\\
Since 2004, Nearctic brown and collared lemmings were live-trapped 3 times during the summer (mid-June, mid-July, and mid-August) in two 11 ha grids. Each grid is made of 144 traps separated by 30 m according to a cartesian plane, one in mesic habitat and the other in wet habitat, located in the Qarlikturvik valley \citep{fauteux2015, gauthier2020lemmings}. Density of each species was estimated at each occasion using spatially explicit capture-recapture methods (see \citet{fauteux2015} for details). From 1995 to 2016 snap-trapping was performed once a year (mid-July) along 2 groups of transects located in the same habitats than the trapping grids \citep{gruyer2008}. Index of abundance derived from snap-trapping were transformed in density estimates in each habitat for the period 1995-2003 using the equation provided by \citet{fauteux2018} based on the period of overlap between the two sampling methods (2004 to 2016). 
\\
\item[] \textit{\textbf{Arctic fox movement tracking}}\\
In order to assess fox abundance based on the size of their home range, 109 Arctic foxes were fitted with Argos Platform Transmitter Terminals mounted on collars between 2008 and 2016 \citep{lai2015,christin2015, dulude2023}. Foxes were captured between May and August across the study area, within and outside the goose colony \citep{dulude2023}. Sampling of animal locations was set for an interval of 1 or 2 days and only locations between May 1 and October 30 were retained \citep{dulude2023}.
\\
\item[] \textit{\textbf{Parasitic jaeger banding}}\\
In 2009, a significant effort was made to band as many parasitic jaegers as possible within the study area. This effort resulted in the banding of 17 adult individuals (Therrien and Gauthier, unpublished data).
\end{enumerate}
\subsubsection*{e. Species body mass}
All vertebrate individuals captured for marking purposes were systematically weighed: snow goose (G. Gauthier, M.-C. Cadieux and J. Lefebvre, unpublished data), snowy owl \citep{therrien2012, robillard2018}, American-golden plovers \citep{lamarre2021}, common-ringed plovers \citep{leandri2019}, other shorebirds (J. Bêty, unpublished data), glaucous gulls \citep{gauthier2015}, long-tailed jaeger \citep{seyer2019}, parasitic jaegers (J.-F. Therrien and G. Gauthier, unpublished data), Lapland longspurs (J. Bêty and G. Gauthier, unpublished data), lemmings \citep{gauthier2020lemmings}, American ermine (Bilodeau and Bolduc, unpublished data) and Arctic foxes \citep{lai2015}. When not available, we extracted mean body mass from the literature \citep{wilman2014}.
\newpage

@@

        \begin{enumerate}[label=\alph*.]       
     \item[] \textit{\textbf{Pacific loon, red-throated loon, cackling goose, tundra swan and glaucous gull}}   \\
Based on the systematic and intensive search for the glaucous gull, cackling goose, tundra swan, red- throated loon and Pacific loon nests in wetlands, we are confident that we have found nearly all nests across the study area from 2017 to 2019 and in 2022. We observed a relatively strong correlation between the nest density of glaucous gulls in the Qarlikturvik valley and the nest density across the entire study area (R\textsuperscript{2}= 0.7, p = 0.16, n= 4). Consequently, we estimated the density of glaucous gulls at the scale of the study area between 2004 and 2016 based on the nest density in the Qarlikturvik valley ($y = 0.12409x + 0.13774$). However, we did not observe such strong relationships for loons and swans and thus we did not extend the time series. Regarding cackling geese, we observed signs of an exponential increase over time based on the annual number of nests found in various zones of the study area. We thus fitted an exponential model using the number of nests found annually over two distinct periods: in 1996 when the first nest was discovered, and then from 2017 to 2023 when sampling effort was systematic across the whole study area (\textbf{Figure \ref{figure:cackling}}). We used the fitted model to estimate abundance between 1996 and 2016 when monitoring was less systematic, which could potentially underestimate observed abundance as seen on \textbf{Figure \ref{figure:cackling}}. We multiplied nest density by two to obtain the abundance (assuming two individuals per nest).

@@

                \item[] \textit{\textbf{Rough-legged hawk, peregrine falcon and snowy owl}} \\
                We estimated the abundance of breeding rough-legged hawks, peregrine falcons and snowy owls based on systematic nest monitoring conducted throughout the study area for these species. To convert the number of nests into breeding abundance, we multiplied it by two (assuming two individuals per nest). For snowy owls, we extended the time series from 1996 to 2011 based on a linear regression between nest density in the Qarlikturvik valley and nearby plateaus (Black and Southern plateaus) and nest density across the entire study area ($y= 0.68867x -0.00173$; R\textsuperscript{2}= 0.98; p < 0.0001, n= 10). We used the same approach for rough-legged hawks ($y=0.49851x$, R\textsuperscript{2}= 0.99, p < 0.0001, n= 8) to extend the time series from 2007 to 2012. We did not extend the time series for peregrine falcons because the correlation is not as strong (R\textsuperscript{2}= 0.20, p= 0.27, n=8).

@@

                \item[] \textit{\textbf{Sandhill crane}}\\
               We determined the mean nest density of sandhill cranes from the nest sampling between 2004 and 2023 on the southern side of the glacial river in the Qarlikturvik valley. We determined mean nest density by dividing the mean annual number of nests recorded by the area of the surveyed zone (33 km\textsuperscript{2}). We estimated total abundance by multiplying mean nest density by the total area of the study area. We assumed a uniform density across the study area, as sandhill crane nests have been observed in wetland, mesic, and upland habitats.

                \item[] \textit{\textbf{American golden-plover and black-bellied plover}}\\
                We used a distance sampling approach to estimate the abundance of American golden-plovers in the lowland zones of the study area between 2014 and 2023. Observations of plovers were made along vertebrate count transects mainly in mesic habitat. Perpendicular distance between detected individuals and the transect path were used (n= 1015) to estimate a detection function with the \textit{ds} function from the \textit{Distance} package \citep{miller2019}. To determine the detection function, we applied a truncation distance of 150 m (i.e., maximum distance on either side of the observer where observations have been considered) and selected the model with the lowest AIC, which included a "hn" key and a single "cos" adjustment term. We excluded observations of groups with more than four individuals, as these likely indicated groups of non-breeders. We did not estimate abundance in wetland habitat because American golden-plovers nest almost exclusively in mesic habitat \citep{parmelee1967}. We estimated the abundance in the upland zones (i.e., plateaus) by applying a correction factor to the abundance in lowland zones. This correction factor was based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}).

@@

                \item[] \textit{\textbf{Lapland longspur and Baird’s sandpiper}}\\
                We estimated the mean abundance of Lapland longspurs in the different lowland zones of the study area based on a relation between nest density and the number of individuals observed per transect (\textbf{Figure \ref{figure:lalo_basa}}). For Baird’s sandpiper, we employed a similar approach, but instead of using the mean number of individuals observed per transect, we used the mean proportion of transects where at least one individual was detected. We made this adjustment because this species was less frequently observed. In this relationship, nest density for these two species came from the intensive nest sampling conducted within the core 2 km\textsuperscript{2} area of the Qarlikturvik valley and observations of individuals from transects carried out in the larger 8 km\textsuperscript{2} area in which the core area was located (\textbf{Figure \ref{figure:qarlikturvik_valley}}). This approach allowed us to incorporate a larger sample size from the transects while focusing on a measure of nest density determined systematically. Transects observations in lowland were then converted into nest density using the regressions equation, and then in total number of individuals by multiplying by the area and a factor 2. We estimated the density of both species in the upland zones by applying a correction factor to the annual mean density in lowland zones. This correction factor was based on the relative abundance ratio between the upland and lowland zones (\textbf{Table \ref{table:table_species_relative_abundance_upland}}). We acknowledge that the relation for Baird’s sandpiper is weaker; however, it offers some refinement compared to assuming a uniform density throughout the study area.

                \begin{figure}[H]
                \begin{subfigure}{0.5\textwidth}
                \includegraphics[width=0.95\linewidth]{figures/lapland_longspur_nest_transect.pdf} 
                \caption{}
                \end{subfigure}
                \begin{subfigure}{0.5\textwidth}
                \includegraphics[width=0.95\linewidth]{figures/bairds_sandpiper_nest_transect.pdf}
                \caption{}
                \end{subfigure}
                \caption{a) The red line shows the linear relation between nest density and the number of individuals observed per transect, described by the equation nest density = 2.0031 × number of individuals per transect. The relation was forced through the origin. Blue dots represent annual values (2014–2019 and 2022–2023) of nest density and mean number of individuals per transect from the core sampled area of the Qarlikturvik Valley. We assumed that the relation between nest density and observations on transects is linear. b) The red line shows the linear relation between nest density and the proportion of transects with at least an individual, described by the equation nest density= 30.9519 x proportion of transects with at least one individual. The relation was forced through the origin. Blue dots represent annual values (2014–2019 and 2022–2023) of nest density and mean proportion of transects with at least an individual from the core sampled area of the Qarlikturvik Valley. We assumed that the relation between nest density and observations on transects is linear.}
                \label{figure:lalo_basa}
                \end{figure}

@@

                \subsubsection*{d. Project personnel}
                \begin{enumerate}[label=\alph*.]
                        \item[] \textit{\textbf{Principal and associated investigators}}\\
                        Gilles Gauthier, Austin Reed, Jean-François Giroux, Dominique Berteaux, Joël Bêty, Josée Lefebvre, Dominique Gravel, Jean-François Therrien, Nicolas Lecomte, Dominique Fauteux, Pierre Legagneux (see \citet{gauthier2024b})
                        \item[] \textit{\textbf{Students}}\\
                        By combining animal and plant ecology, 24 doctoral theses and 56 master theses have been completed in relation to the study area located on the south plain of Bylot Island (see \citet{gauthier2024b}).
                \end{enumerate}
\newpage 
 \section*{\textcolor{Blue}{Class III. Data set status and accessibility}}
    \section*{A. Status}
                \subsection*{1. Latest update} November 8, 2024
        \subsection*{2. Latest archive date} November 8, 2024
        \subsection*{3. Metadata status} November 8, 2024
        \subsection*{4. Data verification} The methods employed to estimate species abundance were subject to several rounds of revision by the authors.
   \section*{B. Accessibility}
                \subsection*{1. Storage location and medium}
The complete data set will be published as supporting information should the manuscript be accepted for publication. Metadata document, raw data and codes will also be archived in Dryad.

@@

                        \subsection*{2. Size} 
                        a. 35 records, not including header row (4.3 kB)\\
                        b. 271 records, not including header row  (35.1 kB)\\
                        c. 53 records, not including header row  (3.7 kB)\\
                        d. 18 records, not including header row  (1.1 kB)

@@

                        \subsection*{7. Authentication procedures}
                        Sums of the numeric columns: \\
                        b. \texttt{BYLOT-species\_abundance.csv}: abundance= 2278553\\
                        c. \texttt{BYLOT-species\_body\_mass.csv}: body\_mass\_g= 49617; sample\_size= 13902\\
                        d. \texttt{BYLOT-interannual\_variation\_nest\_density.csv}: mean\_nest\_density\_km2= 20.447; sd\_nest\_density\_km2= 9.529; sample\_size\_nest\_density\_km2= 186

@@

 \section*{\textcolor{Blue}{Class V. Supplemental descriptors}}
    \section*{A. Data acquisition}
    		\subsection*{1. Data forms or acquisition methods}
   		See Section \textbf{2. Experimental or sampling design} 
   		\subsection*{2. Location of completed data forms}
        Raw data and codes will be archived in Dryad.
        \subsection*{3. Data entry verification procedures}
       	The methods used to extract final species abundance estimates were subject to several rounds of revision by the authors.
